---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Data Dump Version: 25-Sep-2021 09:44 from Wikidata Dump

## Download the dump data

```{r echo=TRUE, eval=FALSE}
###################################################################
# You should remove the single comment out "#".
###################################################################
##Install
#devtools::install_github( "kumeS/AHLodDbs", force = TRUE, dependencies = TRUE)
library(AHLodDbs)

##Create Directory
if(!dir.exists("Dump")){dir.create("Dump")}

##Download from the original dump (latest)
#system("wget -b https://dumps.wikimedia.org/wikidatawiki/entities/latest-truthy.nt.bz2")
#system("mv latest-truthy.nt.bz2 ./Dump/WikidataRDF-25-Sep-2021.nt.bz2")

##Alternative way: Download from Public_Repository_for_AHWikiDataDbs
#fp <- system.file("scripts", "gdrive_download.sh", package="AHLodDbs")

##source
#system(paste0("source ", fp))

##Download the file
#system("gdrive_download 1tS3nngpJzsT7-fOb_lrXZWBwwze_xnmV ./Dump/WikidataRDF-25-Sep-2021.nt.bz2 &")
```

## Decompress the bz2 file

```{r echo=TRUE, eval=FALSE}
setwd("./Dump")
getwd()

##unzip
#system("bunzip2 -k WikidataRDF-25-Sep-2021.nt.bz2")

##rename
#system("mv WikidataRDF-25-Sep-2021.nt Dump.nt")
```

## Pre-Processing on R via sh scripts

```{r echo=TRUE, eval=FALSE}
##extract English labels
#system('grep -e "@en ." ./Dump.nt > ./Label_en.nt')
#system('grep -e " <http://www.w3.org/2000/01/rdf-schema#label> " ./Label_en.nt > ./Label_en_rdfs.nt')
#system('grep -e " <http://www.w3.org/2004/02/skos/core#altLabel> " ./Label_en.nt > ./Label_en_Altlabel.nt')
#system('grep -e " <http://schema.org/description> " ./Label_en.nt > ./Label_en_description.nt')

##extract Japanese labels
#system('grep -e "@ja ." ./Dump.nt > ./Label_ja.nt')
#system('grep -e " <http://www.w3.org/2000/01/rdf-schema#label> " ./Label_ja.nt > ./Label_ja_rdfs.nt')
#system('grep -e " <http://www.w3.org/2004/02/skos/core#altLabel> " ./Label_ja.nt > ./Label_ja_Altlabel.nt')
#system('grep -e " <http://schema.org/description> " ./Label_ja.nt > ./Label_ja_description.nt')

##extract class hierarchy
#system('grep -e " <http://www.wikidata.org/prop/direct/P279> " ./Dump.nt > ./Relation_P279.nt')
#system('grep -e " <http://www.wikidata.org/prop/direct/P31> " ./Dump.nt > ./Relation_P31.nt')
#system('cat ./Relation_P279.nt ./Relation_P31.nt > ./Relation_P279_P31.nt')

##extract property list
#system('grep -e "^<http://www.wikidata.org/prop/direct/P[0-9]" ./Dump.nt > ./PropertyList_01.nt')
#system('grep -e "^<http://www.wikidata.org/entity/P[0-9]" ./Dump.nt > ./PropertyList_02.nt')
#system('cat PropertyList_01.nt PropertyList_02.nt > PropertyList.nt')
#system('rm -rf PropertyList_01.nt PropertyList_02.nt')

#system("grep -v -e '\"@' PropertyList.nt > PropertyList_wo.nt")
#system("grep -e '\"@' PropertyList.nt > PropertyList_w.nt")
#system("grep -e '\"@en .' PropertyList_w.nt > PropertyList_w_en.nt")
#system("grep -e '\"@ja .' PropertyList_w.nt > PropertyList_w_ja.nt")
#system('cat PropertyList_w.nt PropertyList_w_en.nt PropertyList_w_ja.nt > PropertyList_all.nt')
#system("grep -e '\"@' PropertyList_all.nt > PropertyList_labels.nt")
#system('rm -rf PropertyList_wo.nt PropertyList_w.nt PropertyList_w_en.nt PropertyList_w_ja.nt')

##Others
#system("grep -v -e '\"@[a-z]' ./Dump.nt > ./Others_i.nt")
#system('grep -v -e " <http://www.wikidata.org/prop/direct/P31> " ./Others_i.nt > ./Others_ii.nt')
#system('grep -v -e " <http://www.wikidata.org/prop/direct/P279> " ./Others_ii.nt > ./Others_iii.nt')
#system('grep -e "^<http://www.wikidata.org/entity/Q[0-9]" ./Others_iii.nt > ./Others_iiii.nt')

##MeSH descriptor ID (P486)
#system('grep -e " <http://www.wikidata.org/prop/direct/P486> " ./Others_iiii.nt > ./Relation_MeSH_descriptor_ID.nt')

##PubChem CID (P662)
#system('grep -e " <http://www.wikidata.org/prop/direct/P662> " ./Others_iiii.nt > ./Relation_PubChem_CID.nt')

##KEGG ID (P665)
#system('grep -e " <http://www.wikidata.org/prop/direct/P665> " ./Others_iiii.nt > ./Relation_KEGG_ID.nt')

##NCBI taxonomy ID (P685)
#system('grep -e " <http://www.wikidata.org/prop/direct/P685> " ./Others_iiii.nt > ./Relation_NCBI_taxonomy_ID.nt')

##Remove files
#system('rm -rf ./Dump.nt ./Relation_P279.nt ./Relation_P31.nt ./Others_i.nt ./Others_ii.nt ./Others_iii.nt')
#system('ls -ul -h')
```

## Pre-Processing via R scripts

```{r echo=TRUE, eval=FALSE}
setwd("../")
getwd()
dir("./Dump")
library(AHLodDbs)
library(magrittr)
```

## For Relation_P279_P31.nt

```{r echo=TRUE, eval=FALSE}
File_path <- "./Dump/Relation_P279_P31.nt"
Wikidata_PurseNT_ClassHierarchy(File_path)
DFcsv2Rds(File_path, Type="wikidata")
file.remove(paste0(sub(".nt$", "", File_path), "_df.csv"))

if(F){
#Check Data
Dat <- readRDS(paste0(sub(".nt$", "", File_path), "_df.Rds"))

head(Dat)
table(grepl("^wd:Q", Dat$Subject))
table(grepl("^rdf", Dat$Property))
table(Dat$Property)
table(grepl("^wd:Q", Dat$Object))
}

```

## For English Labels

```{r echo=TRUE, eval=FALSE}
##rdfs:label
File_path <- "./Dump/Label_en_rdfs.nt"
Wikidata_PurseNT_Label(File_path, readr=T, tsv=T)
DFtsv2Rds(File_path, Type="wikilabel")

##Altlabel
File_path <- "./Dump/Label_en_Altlabel.nt"
Wikidata_PurseNT_Label(File_path, readr=T, tsv=T)
DFtsv2Rds(File_path, Type="wikilabel")

if(F){
##Check Data
Dat <- readRDS("./Dump/Label_en_rdfs_df.Rds")

head(Dat)
table(grepl("^wd:Q", Dat$Subject))
table(Dat$Property)
table(grepl("^wd:Q", Dat$Object))
table(Dat$OtherInfo)
}

#Bind labels
Label_en_rdfs_df <- readr::read_tsv("./Dump/Label_en_rdfs_df.tsv", col_names = F, show_col_types = F)
Label_en_Altlabel_df <- readr::read_tsv("./Dump/Label_en_Altlabel_df.tsv", col_names = F, show_col_types = F)
#head(Label_en_rdfs_df); head(Label_en_Altlabel_df)

table(Label_en_rdfs_df$OtherInfo)
#table(grepl("\\\\\\\\", Label_en_rdfs_df$Object))
#head(Label_en_rdfs_df$Object[grepl("\\\\\\\\", Label_en_rdfs_df$Object)])
table(Label_en_Altlabel_df$OtherInfo)
#table(grepl("\\\\\\\\", Label_en_Altlabel_df$Object))
#head(Label_en_Altlabel_df$Object[grepl("\\\\\\\\", Label_en_Altlabel_df$Object)])

table(Label_en_rdfs_df$X4)
table(Label_en_Altlabel_df$X4)

X <- 1:1000000
table(grepl("\\\\\\\\", Label_en_rdfs_df$X3[X]))
head(Label_en_rdfs_df$X3[X][grepl("\\\\\\\\", Label_en_rdfs_df$X3[X])])
table(grepl("\\\\", Label_en_rdfs_df$X3[X]))
head(Label_en_rdfs_df$X3[X][grepl("\\\\", Label_en_rdfs_df$X3[X])])


>>>>>>> b55e2115b3493440c16d6bad9cfe565a6455bd53

Label_en_All <- rbind(Label_en_rdfs_df, Label_en_Altlabel_df)
dim(Label_en_All)
saveRDS(Label_en_All, "./Dump/Label_en_All.Rds")


```

## For Japanese Labels

You need to install the `nkf` library via Homebrew if MacOSX.

```{r echo=TRUE, eval=FALSE}
##rdfs:label
File_path <- "./Dump/Label_ja_rdfs.nt"
Wikidata_PurseNT_Label(File_path)

##Convert their unicode (Escape character) to Multi-Byte Character (Japanese)
system(paste0("cat ", sub(".nt$", "_df.csv", File_path), " | sed 's/\\\\u\\(....\\)/\\&#x\\1;/g' | nkf --numchar-input -w > ", sub(".nt$", "_df_nkf.csv", File_path)))
ProcCSV_af_nkf(File_path=sub(".nt$", "_df_nkf.csv", File_path))
file.remove(c("./Dump/Label_ja_rdfs_df_nkf.csv", "./Dump/Label_ja_rdfs_df.csv"))

if(F){
##Check Data
Dat <- readRDS("./Dump/Label_ja_rdfs_df.Rds")

head(Dat)
table(grepl("^wd:Q", Dat$Subject))
table(Dat$Property)
table(grepl("^wd:Q", Dat$Object))
table(Dat$OtherInfo)
}

##Altlabel
File_path <- "./Dump/Label_ja_Altlabel.nt"
Wikidata_PurseNT_Label(File_path)

##Convert their unicode (Escape character) to Multi-Byte Character (Japanese)
system(paste0("cat ", sub(".nt$", "_df.csv", File_path), " | sed 's/\\\\u\\(....\\)/\\&#x\\1;/g' | nkf --numchar-input -w > ", sub(".nt$", "_df_nkf.csv", File_path)))
ProcCSV_af_nkf(File_path=sub(".nt$", "_df_nkf.csv", File_path))
file.remove(c("./Dump/Label_ja_Altlabel_df.csv", "./Dump/Label_ja_Altlabel_df_nkf.csv"))

#Bind labels
Label_ja_rdfs_df <- readRDS("./Dump/Label_ja_rdfs_df.Rds")
Label_ja_Altlabel_df <- readRDS("./Dump/Label_ja_Altlabel_df.Rds")

#head(Label_ja_rdfs_df); head(Label_ja_Altlabel_df)
#table(Label_ja_rdfs_df$OtherInfo)
#table(Label_ja_Altlabel_df$OtherInfo)

Label_ja_All <- rbind(Label_ja_rdfs_df, Label_ja_Altlabel_df)
dim(Label_ja_All)
saveRDS(Label_ja_All, "./Dump/Label_ja_All.Rds")
```

## Relation_KEGG_ID.nt

```{r echo=TRUE, eval=FALSE}
File_path <- "./Dump/Relation_KEGG_ID.nt"
Wikidata_PurseNT_Others(File_path)
DFcsv2Rds(File_path, Type="wikidata")
file.remove(paste0(sub(".nt$", "", File_path), "_df.csv"))
```

## Relation_MeSH_descriptor_ID.nt

```{r echo=TRUE, eval=FALSE}
File_path <- "./Dump/Relation_MeSH_descriptor_ID.nt"
Wikidata_PurseNT_Others(File_path)
DFcsv2Rds(File_path, Type="wikidata")
file.remove(paste0(sub(".nt$", "", File_path), "_df.csv"))
```

## Relation_NCBI_taxonomy_ID.nt

```{r echo=TRUE, eval=FALSE}
File_path <- "./Dump/Relation_NCBI_taxonomy_ID.nt"
Wikidata_PurseNT_Others(File_path)
DFcsv2Rds(File_path, Type="wikidata")
file.remove(paste0(sub(".nt$", "", File_path), "_df.csv"))
```

## Relation_PubChem_CID.nt

```{r echo=TRUE, eval=FALSE}
File_path <- "./Dump/Relation_PubChem_CID.nt"
Wikidata_PurseNT_Others(File_path)
DFcsv2Rds(File_path, Type="wikidata")
file.remove(paste0(sub(".nt$", "", File_path), "_df.csv"))
```




